Bricking up backdoors in Intrusion Detection Systems

### 2019-09-10 ex2/paper/paper.tex
Updated fine-tuning
Actually remove conclusion

@@

We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}. The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly
while decisions that are used for the backdoor are hopefully removed. We develop several variants of our pruning defense with increasing levels of sophistication. These variants are pruning \todo{Remove space in the enumeration}
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples. Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
\end{enumerate}
\begin{figure*}[h]
\includegraphics[width=\textwidth]{pruning_example.pdf}
\caption{Example of a \gls{dt} being pruned. To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf. It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside). The leaf that is going to be cut in each step has a dashed connection to its parent.}
\label{fig:pruningExample}
\end{figure*}

@@

\section{Discussion}
\todo{Max: Merge discussion and conclusions to one section and remove most parts of conclusions?}
From our experiments, we can make three main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.

@@

\section{Conclusions}
%In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}. We used a recent publicly available dataset for network \gls{ad} and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.

With this setup, we explored several mitigation techniques. While, surprisingly, neural network pruning and fine-tuning were uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots  for both the \gls{dl} and \gls{rf} models.

%We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.

%Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors. Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use. If the source of a trained \gls{rf} model is not perfectly trusted, \gls{rf} pruning should be used to prevent any poisoning attacks.

>>>>

We make all our code publicly available at \todo{Insert link to GitHub repo}.

@@

We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}. The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly
while decisions that are used for the backdoor are hopefully removed. We develop several variants of our pruning defense with increasing levels of sophistication. These variants are pruning
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples. Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
\end{enumerate}
\begin{figure*}[h]
\includegraphics[width=\textwidth]{pruning_example.pdf}
\caption{Example of a \gls{dt} being pruned. To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf. It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside). The leaf that is going to be cut in each step has a dashed connection to its parent.}
\label{fig:pruningExample}
\end{figure*}

@@

\section{Discussion}
From our experiments, we can make three main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.

@@

%\section{Conclusions}
%%In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}. We used a recent publicly available dataset for network \gls{ad} and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
%
%With this setup, we explored several mitigation techniques. While, surprisingly, neural network pruning and fine-tuning were uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots  for both the \gls{dl} and \gls{rf} models.
%
%%We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
%
%%Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors. Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use. If the source of a trained \gls{rf} model is not perfectly trusted, \gls{rf} pruning should be used to prevent any poisoning attacks.
